==Process Stallard Data tables
```{r setup,echo=FALSE,include=FALSE}
options( stringsAsFactors=FALSE )

require(lubridate)
require(plyr)
require(reshape2)
require(ggplot2)
```
1
in the attached by Stallard Tables 3-6 -- just need the upper block Date of first exposure and Date of claim In data frames

```{r read-table-fnc}
  projyears=seq(2009,2059,5)
rdtbl<-function(infile){
  cnames<-c("date.first.exposure",projyears,"total","development.factor")
  tbl=read.delim(infile,header=F,sep=" ")
  colnames(tbl)=cnames
  str(tbl)
  print(tbl)
  return(tbl)
}

```

```{r read-tables}
table3=rdtbl("table3.txt")
table4=rdtbl("table4.txt")
table5=rdtbl("table5.txt")
table6=rdtbl("table6.txt")
```
2
Build a set of data frames in which the dates of claims are by single year, not 5-year groups
Basically, let y = the cumulative number of claims, going across
Then build a splinefun of the cumulative number of claims
Interpolate the spline fun to individual years
Go from cumulative back to indiv year counts by subtraction
(Standard demographic stuff)

```{r interpolate}
sdata<-function(df) {
  tbl<-df[,1:12]
    for (cc in 3:12) {
    tbl[,cc]=tbl[,cc]+tbl[cc-1]
  }

  tall=melt(tbl)
  tall=tall[order(tall$date.first.exposure),]
  sdf=data.frame(exposure=tall$date.first.exposure,claims=tall$value,year=projyears[as.numeric(tall$variable)])
  #okay so we have 14 exposure periods of 11 obs each
  projvec=projyears[1]:projyears[11]
  nproj=length(projvec)
  xdf=data.frame(data.first.exposure=character(),year=numeric(),cum.proj=numeric(),inc.proj=numeric())
  for (exp in 1:14) {
    rn=exp*11
    r1=rn-11+1
    sps=splinefun(x=projyears,y=sdf$claims[r1:rn])(projvec)
    #print(sps)
#decumulate
    isps=sps
    isps[2:nproj]=sps[2:nproj]-sps[1:(nproj-1)]
    #print(isps)
              tdf=(data.frame(data.first.exposure=rep(df$date.first.exposure[exp],nproj),year=projvec,
                             cum.proj=sps,inc.proj=isps))
   # g<-ggplot()+geom_line(data=tdf,aes(year,inc.proj))+geom_bar(data=sdf[r1:rn],aes(year,claims))
   # print(g)
    xdf=rbind(xdf,tdf)
  }
    return(xdf)
}
proj3=sdata(table3)
proj4=sdata(table4)
proj5=sdata(table5)
proj6=sdata(table6)
```

3
Then build a function
Given a vector of years of first exposure, injury (4-level injuries here), and current date
Return a data frame
     Variables year of future claim
     Expected number of claims (NOT rounded at this point)
Processing
   Each incoming vector element produces a set of probabilities, going across, based on single year of claim counts from 2 above
   Then add these probabilities across all vector elements
SInce the starting date will be in the middle of a year, you will have to prorate the first year of expected counts accordingly
You might allow for an additional weight vector on input

4
Build an inverter function
Suppose we have a matrix of counts from a trust that is in the same form as the Stallard matrix, but by year of claim
We also have a "from date" and a "to date" for dates of claims
We want to estimate how our trust's mix of claims by "Date of first exposure" compares with the Stallard mix
Simple inverter -- between "from date" and "to date" in each first exposure group, calculate the ratio of our total number of claims divided by the stallard number
Return a frame
        date of first exposure (grouped)
        ratio of us to stallard

5
You can feed these ratios into function 3 as "pseudo counts" to get the runoff curves for out trust's particular mix of first exposures
